Argennon: A Scalable Smart Contract Platform

### 2021-06-20 tex/sections/consensus.tex
Update persistence section
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Add the Argennon consensus protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

\note{It will be a randomized proof of stake which is under development....}

\subsection{Estimating A User's Stake}\label{subsec:estimating-a-user's-stake}

In a proof of stake system the influence of a user in the consensus protocol should be proportional to the amount
of stake the user has in the system. Conventionally in these systems, for estimating a user's stake, we use the
amount of native system tokens the user is holding. Unfortunately, one problem with this approach is that a
strong attacker may be able to obtain a considerable amount of system tokens, for example by borrowing from a
DEFI application, and use this stake to attack the system.

To mitigate this problem, for calculating a user's stake at time step \(t\), instead of using the raw ARG
balance, we use the minimum of a \emph{trust value} the system has calculated for the user and the user's
ARG balance:
\[
    S_{u,t} = \min (B_{u,t}, Trust_{u,t})
\]
Where:
\begin{itemize}
    \item \(S_{u,t}\) is the stake of user \(u\) at time step \(t\).
    \item \(B_{u,t}\) is the ARG balance of user \(u\) at time step \(t\).
    \item \(Trust_{u,t}\) is an estimated trust value for user \(u\) at time step \(t\).
\end{itemize}

The agreement protocol, at time step \(t\), will use \(\sum_{u}S_{u,t}\) to determine the required
number of votes for the confirmation of a block, and we let \(Trust_{u,t} = M_{u,t}\), where \(M_{u,t}\) is the
exponential moving average of the ARG balance of user \(u\) at time step \(t\).

In our system a user who held ARGs and participated in the consensus for a long time is more trusted
than a user with a higher balance whose balance has increased recently. An attacker who has obtained a large
amount of ARGs, also needs to hold them for a long period of time before being able to attack the system.

For calculating the exponential moving average of a user's balance at time step \(t\), we can use the following
recursive formula:
\[
    M_{u,t} = (1 - \alpha) M_{u,t-1} + \alpha B_{u,t} = M_{u,t-1} + \alpha (B_{u,t} - M_{u,t-1})
\]
Where the coefficient \(\alpha\) is a constant smoothing factor between \(0\) and \(1\) which represents the
degree of weighting decrease, A higher \(\alpha\) discounts older observations faster.

Usually an account balance will not change in every time step, and we can use older values of EMA for calculating
\(M_{u,t}\): (In the following equations the \(u\) subscript is dropped for simplicity)
\[
    M_{t} = (1 - \alpha)^{t-k}M_{k} + [1 - (1 - \alpha)^{t - k}]B
\]
Where:
\[
    B = B_{k+1} = B_{k+2} = \dots = B_{t}
\]
We know that when \(|nx| \ll 1\) we can use the binomial approximation \({(1 + x)^n \approx 1 + nx}\). So, we can
further simplify this formula:
\[
    M_{t} = M_{k} + (t - k) \alpha (B - M_{k})
\]

@@

In our system a newly created account will not have voting power for some time, no matter how high its
balance is. While this is a desirable property, in case a large proportion of total system tokens are
transferred to newly created accounts, it can result in too much voting power for older accounts. This may decrease
the degree of decentralization in our system.

However, this situation is easily detectable by comparing the total stake of the system with the total balance of
users. If after confirming a block the total stake of the system goes too low and we have:
\[
    \sum_{u}S_{u,t} < \gamma \sum_{u}B_{u,t}
\]
The protocol will perform a \emph{time shift} in the system: the time step of the system
will be incremented for \(m\) steps while no blocks will be confirmed. This will increase the value of \(M_{u,t}\)
for new accounts with a non-zero balance, giving them more influence in the agreement protocol.

For calculating the value of \(m\) which determines the amount of time shift in the system, we should note that when
\(B_{u,t} = B_{u, t-1} = B_u\), we can derive a simple recursive rule for the stake of a user:
\[
    S_{u,t} = (1 - \alpha) S_{u,t-1} + \alpha B_u
\]
Therefore, we have:
\[
    \sum_{u}S_{u,t} = (1 - \alpha) \sum_{u}S_{u,t - 1} + \alpha \sum_{u}B_u
\]
This equation shows that when the balance of users is not changing over time the total stake of the system is the
exponential average of the total ARGs of the system. Consequently, when we shift the time for \(m\) steps, we can
calculate the new total stake of the system from the following equation:

\[
    \sum_{u}S_{u,t+m} = (1 - \alpha)^{m}\sum_{u}S_{u,t} + [1 - (1 - \alpha)^{m}]\sum_{u}B_u
\]
Hence, if we want to increase the total stake of the system from \(\gamma \sum_{u}B_u\) to \(\lambda \sum_{u}B_u\),
we can obtain \(m\) from the following formula, assuming \(\alpha\) is small enough:
\[
    m = \frac{1}{\alpha} \ln \left(\frac{1 - \gamma}{1 - \lambda}\right)
\]

>>>>

Argennon has two primary certificate committees: the committee of \emph{delegates} and the committee
of \emph{validators}. A block is considered final only after it gets a certificate from both of these committees.

The delegates committee generates a certificate for every block of the Argennon blockchain, and the
validators committee generates a certificate every $m$ blocks. The validators committee will
certify a block only if it has already been certified by the committee of delegates.

In addition to primary committees, Argennon has several community driven committees. Certificates of these
committees are not required for block validation, but they could be used by the members of the validators
committee to better decide about the validity of a block.

When an anomaly is detected in the consensus mechanism, the \emph{recovery} protocol is initiated by nodes. The
recovery protocol is designed to be resilient to many types of attacks in order to be able to restore the normal
system functionality.

\subsection{The Committee of Delegates}\label{subsec:the-committee-of-delegates}

The committee of delegates is a small committee of trusted delegates, elected by Argennon users either through the
Argennon Decentralized Autonomous Governance system (ADAGs\footnote{pronounced \textipa{/eI-dagz/}.}) or by emergency
voting during the recovery protocol. At the start of the Argennon mainnet, this committee will have
five members, and later its size could be changed by the ADAGs in a procedure described
in Section~\ref{sec:governance}.

The committee of delegates is responsible for creating new blocks of the Argennon blockchain, and it issues a
certificate for every block of the Argennon blockchain. A certificate needs to be signed
by \textbf{all} of the committee members in order to be considered valid.

Usually, the delegates are large organizations, and they have enough computational resources to generate blocks
very fast. However, a block is not completely final if it does not have the certificate of the validators committee.
A certified block by the delegates will not be accepted by the network, if the last block certified by
the validators committee is behind more than a certain number of blocks.

The committee of delegates may use any type of agreement protocol to reach consensus on the
next block. Usually a very simple and fast protocol can do the job: one of the members
is randomly chosen as the proposer, and other members vote "yes" or "no" on the proposed block.

If one of the delegates lose its network connectivity, no new blocks can be generated. For this reason,
the delegates should invest on different types of communication infrastructure, to make sure they never lose
connectivity to each other and to the Argennon network.

\subsection{The Committee of Validators}\label{subsec:validators-committee}

The Argennon protocol calculates a stake value for every account, which is an estimate of a user's stake in the
system. This value is measured in ARGs. The committee of validators is composed of
\textbf{all} accounts whose stake value is higher than \texttt{minValidatorsStake} threshold. This
threshold is determined by the ADAGs, but it can never be higher
than $100$ ARGs.

Every member of the validators committee has a status which can be either \texttt{online} or \texttt{offline}.
This status is stored in the ARG smart contract, and it can be changed through a method invocation
from this smart contract. When an account sets its status to \texttt{offline}, it receives a small reward, and
it can not change it back to \texttt{online} for \texttt{statusCoolDown} number of blocks.

\note{When an account changes its status, the change has no effect until the block containing the status
change transaction gets certified by the committee of validators.}

A block certificate issued by some members of the validators committee is considered valid, if according to
the staking database of the previous \textbf{certified block by the validators committee}, we have:
\begin{itemize}
    \item The total stake of \texttt{online} accounts is higher than \texttt{minOnlineStake} percent of the
    total stake of all accounts. This threshold can be changed by the ADAGs, but it can never be lower
    than 70 percent.
    \item All signers of the certificate have \texttt{online} status.
    \item The sum of stake values of the certificate signers is higher than 80 percent of the total stake
    of accounts that have \texttt{online} status.
\end{itemize}

If the total online stake goes lower than \texttt{minOnlineStake} threshold, nodes initiate the recovery
consensus protocol.

As we mentioned before, every $m$ blocks of the Argennon blockchain needs a certificate from the committee
of validators. This means that assuming the last certified block by the validators is valid, for issuing the block
certificate, the committee needs to conditionally validate $m$ blocks. We call these $m$ blocks the
validation set. To reduce resource usage of validators' nodes, the protocol requires honest nodes to only
conditionally validate a subset of the validation set. Validators that have a stake lower than
average should only validate one block in expectation.

More precisely, to decide about signing the block certificate, every honest member of the validators
committee iterates over blocks of the validation set, and selects a block with
probability $p_u$. In this way, a member of the validators committee will select $p_{u}m$ blocks in expectation.
Then, the honest member conditionally validates the selected blocks. If the member has selected more than one block,
he may validate the selected blocks in parallel.
If all checked blocks are valid, the validator issues an "accept"
signature, otherwise he issues a "reject" signature.

One recommended formula for calculating the probability of selecting a block for validation by user $u$ is:
\begin{equation}
    p_u=\max\left(\frac{s_u}{m\mu},\ \frac{1}{m}\right),\label{eq:block-selection}
\end{equation}
where $\mu$ is the average stake of committee members and $s_u$ is the stake of user $u$.

If the committee of validators, at any point issues a "reject" certificate for a block, the network will initiate the
recovery protocol.

\subsubsection{Signature Aggregation}

In Argennon, signature aggregation is mostly performed by ZK-EDB servers. To distribute the aggregation workload
between different servers, The committee of validators is divided into pre-determined groups, and each ZK-EDB
server is responsible for signature aggregation of one group. To make sure that there is enough redundancy, the number
of groups should be less than the number of ZK-EDB servers and each group should be assigned to
multiple ZK-EDB servers.

Any member of a group knows all the servers that are responsible for signature aggregation of his group. When a member
signs a block certificate, he sends his signature to \textbf{all} the severs that aggregate the signatures of his group.
These servers aggregate the signatures they receive, and then send the aggregated signature to the delegates.
Furthermore, the delegates aggregate these signatures to produce the final block certificate,
and then include it in the next block.

The role of the delegates in the signature aggregation is limited. The important part of the work is done by ZK-EDB
servers. As long as there are enough honest ZK-EDB servers, the network will be able to perform signature aggregation
even if the delegates are malicious. (See Section~\ref{subsec:recovery})

\subsection{Estimating A User's Stake}\label{subsec:user's-stake}

In a proof of stake system the influence of a user in the consensus protocol should be proportional to the amount
of stake the user has in the system. Conventionally in these systems, a user's stake is considered to be equal with the
amount of native system tokens, the user has "staked" in the system. A user stakes his tokens by locking them in
his account or a staking account for some period of time, and during this time he will not be able to transfer
his tokens.

Unfortunately, there is a subtle problem with this approach. It is not clear that in a real world economic system
how much of the main currency of the system can be locked and kept out of the circulation indefinitely. It seems that
this amount for currencies like US dollar, is quite low comparing to the total market cap of the currency.
This means that for a real world currency this type of staking mechanisms will result in putting the
fate of the system in the hands of the owners of a small fraction of the total supply of a currency.

To mitigate this problem, Argennon uses a hybrid approach for estimating the stake of a user.
Every \texttt{stakingDuration} blocks, which is called a \emph{staking period}, Argennon calculates
a \emph{trust value} for each user. The user's stake
at time step \(t\), is estimated based on the user's trust value and his ARG balance:
\[
    S_{u,t} = \min (B_{u,t}, Trust_{u,k}),
\]
where:
\begin{itemize}
    \item \(S_{u,t}\) is the stake of user \(u\) at time step \(t\).
    \item \(B_{u,t}\) is the ARG balance of user \(u\) at time step \(t\).
    \item \(Trust_{u,k}\) is an estimated trust value for user \(u\) at staking period \(k\).
\end{itemize}

Argennon users can lock their ARG tokens in their account for any period of time. During this time a user
will not be able to transfer his tokens and there is no way for cancelling a lock.
The trust value of a user is calculated based on the amount of his locked tokens and the
Exponential Moving Average (EMA) of his ARG balance:
\[
    Trust_{u,k} = \max (L_{u,k}, M_{u,t_k}),
\]
where
\begin{itemize}
    \item $L_{u,k}$ is the amount of locked tokens of user $u$, whose release time is \textbf{after the end} of
    the staking period $k+1$.
    \item $M_{u,t_k}$ is the Exponential Moving Average (EMA) of the ARG balance of user \(u\) at time step \(t_k\).
    $t_k$ is the start time of the staking period $k$.
\end{itemize}

In Argennon a user who held ARGs and participated in the consensus for a long time is more trusted
than a user with a higher balance whose balance has increased recently. An attacker who has obtained a large
amount of ARGs, also needs to hold them for a long period of time before being able to attack the system.

For calculating the EMA of a user's balance at time step \(t\), we can use the following
recursive formula:
\[
    M_{u,t} = (1 - \alpha) M_{u,t-1} + \alpha B_{u,t} = M_{u,t-1} + \alpha (B_{u,t} - M_{u,t-1})
\]
where the coefficient \(\alpha\) is a constant smoothing factor between \(0\) and \(1\), which represents the
degree of weighting decrease. A higher \(\alpha\) discounts older observations faster.

Usually an account balance will not change in every time step, and we can use older values of EMA for calculating
\(M_{u,t}\): (In the following equations the \(u\) subscript is dropped for simplicity)
\[
    M_{t} = (1 - \alpha)^{t-k}M_{k} + [1 - (1 - \alpha)^{t - k}]B
\]
where:
\[
    B = B_{k+1} = B_{k+2} = \dots = B_{t}
\]
We know that when \(|nx| \ll 1\) we can use the binomial approximation \({(1 + x)^n \approx 1 + nx}\). So, we can
further simplify this formula:
\[
    M_{t} = M_{k} + (t - k) \alpha (B - M_{k})
\]

@@

\subsection{The Recovery Protocol}\label{subsec:recovery}

\note{not yet written...}
